---
title: "Class17 Vaccination Mini Project"
author: "Eduardo Modolo"
format: pdf 
---

```{r}
# Import vaccination data
vax <- read.csv( "covid19vaccinesbyzipcode_test.csv" )
head(vax)
```

>Q1. What column details the total number of people fully vaccinated ?

vax$persons_fully_vaccinated 

>Q2. What column details the Zip code tabulation area?

vax$zip_code_tabulation_area

>Q3. What is the earliest date in this dataset ?

```{r}
head(sort(vax$as_of_date))
```

earliest is Jan 5th 2021

>Q4. What is the latest date in this dataset?

```{r}
head(sort(vax$as_of_date, decreasing = TRUE))
```

latest is feb 28th 2023

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset?

13 numeric columns

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```
you can also look at the skim table

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
length(vax$persons_fully_vaccinated)
```
```{r}
round(sum(is.na(vax$persons_fully_vaccinated)) / length(vax$persons_fully_vaccinated) * 100, 2)
```

>Q8. [Optional]: Why might this data be missing?

some groups of people refuse to give vaccination information, such as military bases and stuff

##Working with dates

```{r}
library(lubridate)
today()
```
```{r}
# This will give an Error!
#today() - vax$as_of_date[1]
```

```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```
how old I am in days
```{r}
today() - ymd("2000-11-05")
```
how old I am in seconds
```{r}
24*60*60*8157
```
How old Avery is in days

```{r}
today() - ymd("2001-07-28")
```
>Q9. How many days have passed since the last update of the dataset?

```{r}
today() - vax$as_of_date[nrow(vax)]
```
>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```
```{r}
library(dplyr)
n_distinct(vax$as_of_date)
```

##Working With ZIP Codes

```{r}
library(zipcodeR)
geocode_zip('92037')
```
```{r}
zip_distance('92037','92109')
```
```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
zip_distance('92037','95765')
```

```{r}
geocode_zip('95765')
```

#Focus on the San Diego area

```{r}
# Subset to San Diego county only areas
sd <- vax[ vax$county == "San Diego" , ]
```

```{r}
head(sd)
nrow(sd)
```

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

Use dplyr to filter for population > 10,000
```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
nrow(sd.10)
```


>Q11. How many distinct zip codes are listed for San Diego County?

```{r}
n_distinct(sd$zip_code_tabulation_area)
```

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]
```
>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of last entry "2023-02-28”?

```{r}
 x <- filter(sd, as_of_date == "2023-02-28")
head(x)
```

```{r}
mean(x$percent_of_population_fully_vaccinated, na.rm = TRUE) * 100
```
average percentage is 74.01%

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2023-02-28”?

```{r}
library(ggplot2)
ggplot(x, aes(percent_of_population_fully_vaccinated)) + geom_histogram(bins = 10, na.rm= T, color= "black", fill = "red") + ggtitle("Histogram of Vaccination rates across San Diego County") + xlab("Percent of Population Fully Vaccinated in a Zip Code Area") + ylab("Count(Zip Code Areas")
```

##Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(title = "Vaccination Rate for La Jolla CA 92037", x= "Date", y="Percent Vaccinated") 
```

##Comparing to similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2023-02-28")

head(vax.36)
```

>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-02-28”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```


```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(title = "Vaccination Rate for La Jolla CA 92037", x= "Date", y="Percent Vaccinated") + geom_hline(yintercept = 0.7213331, color = "red", linetype = 5)
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-02-28”?


```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

>Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +aes(percent_of_population_fully_vaccinated) +geom_histogram(col= "white", fill = "Black", na.rm = TRUE) + xlim(0,1) 
```


>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
x <- filter(vax.36, zip_code_tabulation_area %in% c("92109", "92040"))
x$percent_of_population_fully_vaccinated
```

below the mean of 0.7213

>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="Blue", na.rm = TRUE) +
  ylim(0,1) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with population above 36k are shown") +
  geom_hline(yintercept = 0.7213331, linetype=5)
```





